---
title: "states data"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
```{r warning=FALSE, message=FALSE} 
load(file="states_data.RData")
```

```{r}
library(tidyverse) 
states_data
```
Dataset consists of observations on 50 States plus District of Columbia on the following measures:

## columns in the dataset: 
__`state`__: name of US State

### vote totals presidential elections

Source: MIT Election Data and Science Lab, "County Presidential Election Returns 2000-2020", <https://doi.org/10.7910/DVN/VOQCHQ>  

__`trump_votes_2016`__: each the vote totals for each candidate or all others combined.  

__`clinton_votes_2016`__:

__`other_votes_2016`__  :  

__`trump_votes_2020`__   :               

__`biden_votes_2020`__   :               

__`other_votes_2020`__ :

### COVID cases and deaths

Source: The New York Times. (2021). <https://github.com/nytimes/covid-19-data>

__`cases_2020`__ : COVID total cases for 2020

__`cases_jan_2020`__ to __`cases_oct_2021`__ : cumulative COVID case counts by month

__`vax_dec_2020`__ to __`vax_oct_2021`__ : cumulative vaccination counts per month (all ages)

__`deaths_jan_2020`__ to __`deaths_oct_2021`__: cumulative COVID attributed deaths by month  

__`deaths_pc_jun_2020`__: per capita (10,000 people) COVID deaths by June 2020


### COVID shutdown

Source: Ballotopedia.org.
<https://ballotpedia.org/States_that_issued_lockdown_and_stay-at-home_orders_in_response_to_the_coronavirus_(COVID-19)_pandemic,_2020>

__`stay_begin`__: date of beginning stay at home order 

__`stay_end`__: date of ending of stay at home order

__`shutdown_date_range`__: date length of State stay at home order

__`shutdown_length`__: length of State stay at home order in numeric number of days

### demographics 

US Census Bureau, <https://www.census.gov/data.html> And American Community Survey 5-year estimate 2018. 

__`total_population_2010`__ : State population from 2010 decennial US Census, US Census Bureau

__`total_population_est_2019`__  : estimated population in 2019, 1-year estimate, US Census Bureau.  

__`census_fips_code`__: numeric code for State

The following are American Community Survey 5-year estimates 2018. population totals and by first racial group mention, income, and educational attainment. And Gini coefficient.

__`pop`__: total population

__`white_pop`__:

__`black_pop`__: 

__`native_pop`__: 

__`asian_pop`__: 

__`latino_pop`__: 

__`median_income`__: In US Dollars

__`bachelors_pop`__: Total population 25 or older with a bachelor's degree

__`gini`__:  Gini coefficient of wealth inequality, 2018 estimate

### economics

Source: Bureau of Economic Analysis. Annual GDP, personal income, and employment by State: <https://apps.bea.gov/itable/iTable.cfm?ReqID=70&step=1>.   

__`q1_gdp_2020`__ :   percentage change in GDP from Q4 2019 to Q1 2020 

__`q2_gdp_2020`__  :  percentage change in GDP from Q1 2020 to Q2 2020  

__`q3_gdp_2020`__   : percentage change in GDP from Q2 2020 to Q3 2020   

__`q4_gdp_2020`__    :  percentage change in GDP from Q3 2020 to Q4 2020 

__`pc_income_2020`__ : State per capita income in 2020                    
             
__`oct_2020_unemployed`__ : proportion of population unemployed in October 2020               


### public health

__`drug_alc_deaths_2019`__ : number of deaths attributed to drug or alcohol overdose in 2019. Source: Centers for Disease Control and Prevention, National Center for Health Statistics.   <http://wonder.cdc.gov/ucd-icd10.html>  

__`life_expec`__: life expectancy at birth in years, 2018 Source: KFF's State Health Facts.  Life Expectancy at Birth (in years). Kaiser Family Foundation. <https://www.kff.org/other/state-indicator/life-expectancy/>


### policy liberalism and public opinion mass liberalism

Source: Caughey, Devin; Warshaw, Christopher, 2017, "caughey_warshaw_summary.tab", Replication Data for: Policy Preferences and Policy Change: Dynamic Responsiveness in the American States, 1936–2014, <https://doi.org/10.7910/DVN/K3QWZW/5DVZFO>, Harvard Dataverse, V2, UNF:6:gclIUlb4vZOuKOkJEYoI9g== [fileUNF]  
__`state_social_liberalism_policy`__: index of social policy liberalism 2014, higher scores more liberal

__`state_econ_liberalism_policy`__:  index of economic policy liberalism 2014, higher scores more liberal     

__`mass_social_liberalism_policy`__: index of mass social liberalism 2014, higher scores more liberal  

__`mass_econ_liberalism_policy`__: index of mass economic liberalism 2014, higher scores more liberal    

__`dempid_est`__: Estimated proportion of Democratic Party identifiers in the State voting age public, 2014.

__`reppid_est`__: Estimated proportion of Republican Party identifiers in the State voting age public, 2014.

### legislature characteristics

__`percent_women_legislature`__: percent of women in the State legislature. Source: Center for American Women and Politics, Eagleton Institute of Politics, & Rutgers, The State University of New Jersey (Eds.). (2021, June 24). Women in State Legislatures 2021. Cawp. <https://cawp.rutgers.edu/women-state-legislature-2021>. 

 
